---
title: "Exploratory Analysis"
author: "Marc Wettengel"
date: "06/01/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE)

```


```{r, include = FALSE}
library(ggplot2)
library(gghighlight)
library(dplyr)
library(ggpubr)
library(cowplot)
library(MASS)
library(psych)
Monthly_avg <- read.csv("Monthly_avg.csv")
Data_complete <- read.csv("Data_complete.csv")
Epiweek_avg <- read.csv("Epiweek_avg.csv")
Epiweek_avg <- Epiweek_avg[c(2,3,4,5,6,7,8,9,10,12,13,14,15,16,17)]

Monthly_avg$Month <- as.factor(Monthly_avg$Month)
Data_complete$Month <- as.factor(Data_complete$Month)

Monthly_avg$Year <- as.factor(Monthly_avg$Year)
Data_complete$Year <- as.factor(Data_complete$Year)
```

```{r, include = FALSE}
#Data_april_out <- subset(Data_complete, Month == 4 & outbreak == 1)
#Data_april_non <- subset(Data_complete, Month == 4 & outbreak == 0)
#Data_mar_out <- subset(Data_complete, Month == 3 & outbreak == 1)
#Data_mar_non <- subset(Data_complete, Month == 3 & outbreak == 0)
#Data_spring <- rbind(Data_april_out, Data_mar_out, Data_april_non, Data_mar_non)
#Data_spring_non <- rbind(Data_april_non, Data_mar_non)

#Data_nov_out <- subset(Data_complete, Month == 11 & outbreak == 1)
#Data_nov_non <- subset(Data_complete, Month == 11 & outbreak == 0)
#Data_dec_out <- subset(Data_complete, Month == 12 & outbreak == 1)
#Data_dec_non <- subset(Data_complete, Month == 12 & outbreak == 0)
#Data_fall_out <- rbind(Data_nov_out, Data_dec_out)
#Data_fall_non <- rbind(Data_nov_non, Data_dec_non)
```

- Would it make sense to include the years 2020 and 2021 in the exploratory phase bust not with the stat analysis? Logistic regression it could be included (outbreak period), however, poisson should not have these years (pandemic affected the case counts). 



## Rainfall
```{r}
# plot the monthly averages trends for each year.
ggplot(data=Monthly_avg, aes(x=Month, y=Total_Rain, group = Year, colour = Year)) + 
  geom_line() + 
  labs(y= "Rainfall (mm)", x = "Month") +
  ggtitle("Total Monthly Rainfall") +
  scale_x_discrete(label = month.abb)

Epiweek_avg$Year <- as.factor(Epiweek_avg$Year)
Epiweek_avg$epiweek <- as.numeric(Epiweek_avg$epiweek)

ggplot(data=Epiweek_avg, aes(x=epiweek, y=Total_Rain, group = Year, colour = Year)) + 
  geom_line() +
  theme(
    axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")) +
  labs(y= "Rainfall (mm)", x = "Epiweek") +
  ggtitle("Average Rainfall") +
  scale_x_continuous(breaks=seq(0,53,5))



ggplot(data=Epiweek_avg, aes(x=epiweek, y=Temperature, group = Year, colour = Year)) + 
  geom_line() +
  theme(
    axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")) +
  labs(y= "Temperature (C)", x = "Epiweek") +
  ggtitle("Average Sea Surface Temperature") +
  scale_x_continuous(breaks=seq(0,53,5))



ggplot(Data_complete, aes(x = Month, y = Total_Rain)) +
  geom_boxplot(position = "dodge") + 
  labs(y= "Rainfall (mm)", x = "Month")

ggplot(Data_complete, aes(x = Month, y = sqrt(Total_Rain))) +
  geom_boxplot(position = "dodge") + 
  labs(y= "square root of Rainfall (mm)", x = "Month")

ggplot(Data_complete, aes(x = Month, y = Total_Rain, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw()+
  labs(y= "Rainfall (mm)", x = "Month")

ggplot(Data_complete, aes(x = Month, y = sqrt(Total_Rain), fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw()+
  labs(y= "Rainfall (mm)", x = "Month")

ggplot(Data_complete, aes(x = Month, y = sqrt(Total_Rain), fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(y= "Rainfall (mm)", x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Square Root of Rainfall") +
  scale_color_manual(labels = c("Non Outbreak", "Outbreak"))

ggplot(Data_complete, aes(x = Month, y = sqrt(Total_Rain), fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(y= "Rainfall (mm)", x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Square Root of Rainfall") +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"))
```

### Univariate analysis
#### April
```{r}
#wilcox.test(Data_april_non$Salinity, Data_april_out$Salinity )
#t.test(Data_april_non$Salinity, Data_april_out$Salinity)
```
#### March
```{r}
#wilcox.test(Data_mar_non$Salinity, Data_mar_out$Salinity)
#t.test(Data_mar_non$Salinity, Data_mar_out$Salinity)
```
#### March & April 
```{r}
#wilcox.test(Data_spring_non$Total_Rain, Data_spring_out$Total_Rain)
#t.test(Data_spring_non$Total_Rain, Data_spring_out$Total_Rain)
```
#### November
```{r}
#wilcox.test(Data_nov_non$Total_Rain, Data_nov_out$Total_Rain)
#t.test(Data_nov_non$Total_Rain, Data_nov_out$Total_Rain)
```

## Ocean Salinity
```{r}
# plot the monthly averages trends for each year.
ggplot(data=Monthly_avg, aes(x=Month, y=Salinity, group = Year, colour = Year)) + 
  geom_line() + 
  labs(y= "Salinity (PSU)", x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Monthly Average Sea Surface Salinity")

ggplot(Data_complete, aes(x = Month, y = Salinity)) +
  geom_boxplot(position = "dodge") + 
  labs(y= "Salinity (PSU)", x = "Month")


ggplot(Data_complete, aes(x = Month, y = Salinity, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw()+
  labs(y= "Salinity (PSU)", x = "Month")

ggplot(data=Epiweek_avg, aes(x=epiweek, y=Salinity, group = Year, colour = Year)) + 
  geom_line() +
  theme(
    axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")) +
  labs(y= "PSU", x = "Epiweek") +
  ggtitle("Average Salinity") +
  scale_x_continuous(breaks=seq(0,53,5))

ggplot(Data_complete, aes(x = Month, y = Salinity, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(y= "PSU", x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Average Salinity - Outbreak vs. Non-outbreak") +
  scale_color_discrete(labels = c("Non Outbreak", "Outbreak"))
```

### Univariate analysis
#### November
```{r}
#wilcox.test(Data_nov_non$Salinity, Data_nov_out$Salinity)
#t.test(Data_nov_non$Salinity, Data_nov_out$Salinity)
```
#### December
```{r}
#wilcox.test(Data_dec_non$Salinity, Data_dec_out$Salinity)
#t.test(Data_dec_non$Salinity, Data_dec_out$Salinity)
```
#### November & December
```{r}
#wilcox.test(Data_fall_non$Salinity, Data_fall_out$Salinity)
#t.test(Data_fall_non$Salinity, Data_fall_out$Salinity)
```


## Sea Surface Temperature 
```{r}
lagdata <- read.csv("laggeddata.csv")
# plot the monthly averages trends for each year.
ggplot(data=Monthly_avg, aes(x=Month, y=Temperature, group = Year, colour = Year)) + 
  geom_line() + 
  labs(y= "Temperature (C)", x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Monthly Average Sea Surface Temperature")

ggplot(Data_complete, aes(x = Month, y = Temperature)) +
  geom_boxplot(position = "dodge") + 
  labs(y= "Temperature (C)", x = "Month")


ggplot(Data_complete, aes(x = Month, y = Temperature, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(y= "Temperature (C)", x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Sea Surface Temperature - Calendar Year") +
  scale_color_discrete(labels = c("Non Outbreak", "Outbreak"))

lagdata2 <- na.omit(lagdata)
lagdata3 <- lagdata2
lagdata3$yearid <- as.factor(lagdata3$yearid)

ggplot(data=lagdata3, aes(x=week, y=Total_Rain, group = yearid, colour = yearid)) + 
  geom_line() +
  theme(
    axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")) +
  labs(y= "Rainfall (mm)", x = "Epiweek") +
  ggtitle("Average Rainfall") +
  scale_x_continuous(breaks=seq(0,53,5))
  


#<- ggplot(Epiweek_avg, aes(x = study_week, y = Temperature, color = factor(Outbreak))) +
#  geom_path(aes(group=1), linewidth = 1.2) +
#  theme_classic() +
#  theme(axis.line = element_line(colour = 'black', size = 1.5), 
#        axis.ticks = element_line(colour = "black", size = 1.5),
#        axis.text=element_text(size=20),
#        axis.title=element_text(size=25),
#        legend.title=element_blank(),
#        legend.key.size = unit(2, 'cm'),
#        legend.text = element_text(size=25))+
#  ylab("Sea Surface Temperature (C)")+
#  xlab("Study Week") +
#  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
#             linetype = "dotted", color = "blue", linewidth = 1.4) +
#  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red")) 

```

### Univariate analysis
#### April
```{r}
#wilcox.test(Data_april_non$Temperature, Data_april_out$Temperature)
#t.test(Data_april_non$Temperature, Data_april_out$Temperature)
```
#### March
```{r}
#wilcox.test(Data_mar_non$Temperature, Data_mar_out$Temperature)
#t.test(Data_mar_non$Temperature, Data_mar_out$Temperature)
```
#### March & April 
```{r}
#wilcox.test(Data_spring_non$Temperature, Data_spring_out$Temperature)
#t.test(Data_spring_non$Temperature, Data_spring_out$Temperature)
```
#### November
```{r}
#wilcox.test(Data_nov_non$Temperature, Data_nov_out$Temperature)
#t.test(Data_nov_non$Temperature, Data_nov_out$Temperature)
```

## Photosynthetically Active Radiation 
```{r}
# plot the monthly averages trends for each year.
ggplot(data=Monthly_avg, aes(x=Month, y=PAR, group = Year, colour = Year)) + 
  geom_line() + 
  labs(y= "Monthly average PAR (?)", x = "Month")

ggplot(Data_complete, aes(x = Month, y = PAR)) +
  geom_boxplot(position = "dodge") + 
  labs(y= "PAR (?)", x = "Month")


ggplot(Data_complete, aes(x = Month, y = PAR, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw()+
  labs(y= "PAR (?)", x = "Month")

ggplot(data=Epiweek_avg, aes(x=epiweek, y=PAR, group = Year, colour = Year)) + 
  geom_line() +
  theme(
    axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")) +
  labs(y = expression(paste("Einstein ", m^{2}, " per day")),
       x = "Epiweek") +
  ggtitle("Average Photosynthetically Available Radiation") +
  scale_x_continuous(breaks=seq(0,53,5))

ggplot(Data_complete, aes(x = Month, y = Temperature, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(y = expression(paste("Einstein ", m^{2}, " per day")),
       x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Average PAR - Outbreak vs. Non-outbreak") +
  scale_color_discrete(labels = c("Non Outbreak", "Outbreak"))



```

### Univariate analysis
### April 
```{r}
#wilcox.test(Data_april_non$PAR, Data_april_out$PAR)
#t.test(Data_april_non$PAR, Data_april_out$PAR)
```
#### March
```{r}
#wilcox.test(Data_mar_non$PAR, Data_mar_out$PAR)
#t.test(Data_mar_non$PAR, Data_mar_out$PAR)
```
#### March & April
```{r}
#wilcox.test(Data_spring_non$PAR, Data_spring_out$PAR)
#t.test(Data_spring_non$PAR, Data_spring_out$PAR)
```
#### November
```{r}
#wilcox.test(Data_nov_non$PAR, Data_nov_out$PAR)
#t.test(Data_nov_non$PAR, Data_nov_out$PAR)
```
#### December
```{r}
#wilcox.test(Data_dec_non$PAR, Data_dec_out$PAR)
#t.test(Data_dec_non$PAR, Data_dec_out$PAR)
```
#### November & Decmeber
```{r}
#wilcox.test(Data_fall_non$PAR, Data_fall_out$PAR)
#t.test(Data_fall_non$PAR, Data_fall_out$PAR)
```

## Wind 
```{r}
# plot the monthly averages trends for each year.
ggplot(data=Monthly_avg, aes(x=Month, y=WSPD, group = Year, colour = Year)) + 
  geom_line() + 
  labs(y= "Monthly average wind (m/s)", x = "Month")

ggplot(Data_complete, aes(x = Month, y = WSPD)) +
  geom_boxplot(position = "dodge") + 
  labs(y= "Wind (m/s)", x = "Month")


ggplot(Data_complete, aes(x = Month, y = WSPD, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw()+
  labs(y= "Wind (m/s)", x = "Month")

ggplot(data=Epiweek_avg, aes(x=epiweek, y=WSPD, group = Year, colour = Year)) + 
  geom_line() +
  theme(
    axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")) +
  labs(y= "m/s", x = "Epiweek") +
  ggtitle("Average Wind Speed") +
  scale_x_continuous(breaks=seq(0,53,5))

ggplot(Data_complete, aes(x = Month, y = WSPD, fill = factor(outbreak))) +
  geom_boxplot(position = "dodge") + 
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(y= "m/s", x = "Month") +
  scale_x_discrete(label = month.abb) +
  ggtitle("Average Wind Speed - Outbreak vs. Non-outbreak") +
  scale_color_discrete(labels = c("Non Outbreak", "Outbreak"))
```

### Univariate analysis
#### April
```{r}
#wilcox.test(Data_april_non$WSPD, Data_april_out$WSPD)
#t.test(Data_april_non$WSPD, Data_april_out$WSPD)
```
#### November
```{r}
#wilcox.test(Data_nov_non$WSPD, Data_nov_out$WSPD)
#t.test(Data_nov_non$WSPD, Data_nov_out$WSPD)
```

## Case Count
```{r}
ggplot(data=Epiweek_avg, aes(x=epiweek, y=NVcount, group = Year, colour = Year)) + 
  geom_line() +
  theme(
    axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")) +
  labs(y= "Number of Cases", x = "Epiweek") +
  ggtitle("Community Norovirus Case Count") +
  scale_x_continuous(breaks=seq(0,53,5))


```


## Time series plots
```{r, include = FALSE}
# Add in the study week
Epiweek_avg$study_week<-1:nrow(Epiweek_avg)
Epiweek_avg_test <- Epiweek_avg
Epiweek_avg <- Epiweek_avg[,c(1,2,16,6,15,3,4,5,7,8,9,10,11,12,13,14)]
Epiweek_avg$Outbreak <- ceiling(Epiweek_avg$Outbreak)
Epiweek_avg$wdir <- ifelse(Epiweek_avg$WDIR < 0, 360 + Epiweek_avg$WDIR, Epiweek_avg$WDIR)
# Fill in the missing value for PAR using par specific data set
# Epiweek_avg$PAR[is.na(Epiweek_avg$PAR)] <- 4.11
# Epiweek_avg <- Epiweek_avg[,c(1,2,3,4,5,6,7,8,9,10,12,13,14,15,16,17)]

```

```{r, fig.width=20, fig.height=20, echo=FALSE}

#library(cowplot)

p1 <- ggplot(Epiweek_avg, aes(x = study_week, y = NVcount, color = factor(Outbreak))) +
  geom_path(aes(group=1)) +
  theme_classic() +
  theme(axis.title.x=element_blank(), axis.line = element_line(colour = 'black', size = 1.5), 
        axis.ticks = element_line(colour = "black", size = 1.5), 
        axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        legend.title=element_blank(),
        legend.key.size = unit(2, 'cm'),
        legend.text = element_text(size=20),
        legend.position = "top") +
  ylab("Community Case Count")+
  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
             linetype = "dotted", color = "blue", linewidth = 1) +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red"))

p2 <- ggplot(Epiweek_avg, aes(x = study_week, y = Total_Rain, color = factor(Outbreak))) +
  geom_path(aes(group=1)) +
  theme_classic() +
  theme(axis.title.x=element_blank(), axis.line = element_line(colour = 'black', size = 1.5), 
        axis.ticks = element_line(colour = "black", size = 1.5), 
        axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        legend.title=element_blank(),
        legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=10),
        legend.position = "none")+
  ylab("Rainfall (mm)")+
  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
             linetype = "dotted", color = "blue", linewidth = 1) +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red"))

p3 <- ggplot(Epiweek_avg, aes(x = study_week, y = Salinity, color = factor(Outbreak))) +
  geom_path(aes(group=1)) +
  theme_classic() +
  theme(axis.title.x=element_blank(), axis.line = element_line(colour = 'black', size = 1.5), 
        axis.ticks = element_line(colour = "black", size = 1.5), 
        axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        legend.title=element_blank(),
        legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=10),
        legend.position = "none")+
  ylab("Salinity (PSU)")+
  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
             linetype = "dotted", color = "blue", linewidth = 1) +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red"))

p4 <- ggplot(Epiweek_avg, aes(x = study_week, y = Temperature, color = factor(Outbreak))) +
  geom_path(aes(group=1)) +
  theme_classic() +
  theme(axis.title.x=element_blank(), axis.line = element_line(colour = 'black', size = 1.5), 
        axis.ticks = element_line(colour = "black", size = 1.5), 
        axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        legend.title=element_blank(),
        legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=10),
        legend.position = "none")+
  ylab("Sea Surface Temperature (C)")+
  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
             linetype = "dotted", color = "blue", linewidth = 1) +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red")) 

p5 <- ggplot(Epiweek_avg, aes(x = study_week, y = PAR, color = factor(Outbreak))) +
  geom_path(aes(group=1)) +
  theme_classic() +
  theme(axis.title.x=element_blank(), axis.line = element_line(colour = 'black', size = 1.5), 
        axis.ticks = element_line(colour = "black", size = 1.5), 
        axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        legend.title=element_blank(),
        legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=10),
        legend.position = "none")+
  ylab("PAR (Einstein)")+
  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
             linetype = "dotted", color = "blue", linewidth = 1) +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red"))

p6 <- ggplot(Epiweek_avg, aes(x = study_week, y = wdir, color = factor(Outbreak))) +
  geom_path(aes(group=1)) +
  theme_classic() +
  theme(axis.title.x=element_blank(), axis.line = element_line(colour = 'black', size = 1.5), 
        axis.ticks = element_line(colour = "black", size = 1.5), 
        axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        legend.title=element_blank(),
        legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=10),
        legend.position = "none")+
  ylab("Wind Direction (degrees)")+
  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
             linetype = "dotted", color = "blue", linewidth = 1) +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red"))

p7 <- ggplot(Epiweek_avg, aes(x = study_week, y = WSPD, color = factor(Outbreak))) +
  geom_path(aes(group=1)) +
  theme_classic() +
  theme(axis.line = element_line(colour = 'black', size = 1.5), 
        axis.ticks = element_line(colour = "black", size = 1.5),
        axis.text = element_text(size=12),
        axis.title.y=element_text(size=12),
        axis.title.x=element_text(size=20),
        legend.title=element_blank(),
        legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=10),
        legend.position = "none")+
  ylab("Wind Speed (m/s)")+
  xlab("Study Week") +
  geom_vline(xintercept=c(54, 62, 106, 215, 228, 580, 582, 728, 747, 797, 802), 
             linetype = "dotted", color = "blue", linewidth = 1) +
  scale_colour_manual(name = "Outbreak", labels = c("Non Outbreak", "Outbreak"), values = c("black", "red"))

pp <- list(p1, p2, p3, p4, p5, p7)
plot_grid(plotlist=pp, ncol=1, align='v')
```


```{r, fig.width=20, fig.height=22, echo=FALSE}

p11 <- ggplot(Epiweek_avg, aes(x = epiweek, y = NVcount, group = Year, color = factor(Outbreak))) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x=element_blank())+
  ylab("NV Count")+
  scale_colour_manual(values = c("grey", "red"))

p12 <- ggplot(Epiweek_avg, aes(x = epiweek, y = Total_Rain, group = Year, color = factor(Outbreak))) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x=element_blank())+
  ylab("Daily avg. Rainfall")+
  scale_colour_manual(values = c("grey", "red"))

p13 <- ggplot(Epiweek_avg, aes(x = epiweek, y = Salinity, group = Year, color = factor(Outbreak))) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x=element_blank())+
  ylab("Avg. Salinity")+
  scale_colour_manual(values = c("grey", "red"))

p14 <- ggplot(Epiweek_avg, aes(x = epiweek, y = Temperature, group = Year, color = factor(Outbreak))) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x=element_blank())+
  ylab("Avg. SST")+
  scale_colour_manual(values = c("grey", "red"))

p15 <- ggplot(Epiweek_avg, aes(x = epiweek, y = PAR, group = Year, color = factor(Outbreak))) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x=element_blank())+
  ylab("Avg. PAR")+
  scale_colour_manual(values = c("grey", "red"))

p16 <- ggplot(Epiweek_avg, aes(x = epiweek, y = WSPD, group = Year, color = factor(Outbreak))) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x=element_blank())+
  ylab("Avg. WSPD")+
  scale_colour_manual(values = c("grey", "red"))

p17 <- ggplot(Epiweek_avg, aes(x = epiweek, y = WDIR, group = Year, color = factor(Outbreak))) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x=element_blank())+
  ylab("Avg. WDIR")+
  scale_colour_manual(values = c("grey", "red"))

pp1 <- list(p11, p12, p13, p14, p15, p16, p17)
plot_grid(plotlist=pp1, ncol=1, align='v')
```


### DLNM Package
```{r}
library(dlnm)
library(splines)
#vignette("dlnmOverview") #for more information on the package
#testdlnm <- crossbasis(Epiweek_avg$Total_Rain, lag = 12) 

# Assuming Rainfall has a linear relationship
# lag structure is using unconstrained function 
cb1.rain <- crossbasis(Epiweek_avg$Total_Rain, lag = 12, 
                       argvar=list(type="lin"), 
                       arglag=list(type="integer"))

summary(cb1.rain)

model1 <- glm(Outbreak ~ cb1.rain , family="binomial", Epiweek_avg)

pred.cb1.rain <- crosspred(cb1.rain, model1, from = 0, to = 30, bylag = 1, cumul = T)

plot(pred.cb1.rain, "slices", var = 1, col = 3, ci.arg=list(density=15,lwd=2), ci.level=0.80)
plot(pred.cb1.rain, "slices", var = 1, ci = "bars", type="p", col=2, pch=19,
    ci.level=0.80)
plot(pred.cb1.rain, "slices", var = 1, col = 3, cumul = TRUE)

plot(pred.cb1.rain, xlab="Rainfall", zlab="RR", theta=300, phi=30, lphi=30,
    main="3D graph of rainfall effect")

plot(pred.cb1.rain, "contour", xlab="Rainfall", key.title=title("RR"),
    plot.title=title("Contour plot",xlab="Rainfall",ylab="Lag")) 
```

```{r}
cb2.rain <- crossbasis(Epiweek_avg$Total_Rain, lag = 12, 
                       argvar=list(type="thr", thr = 5), 
                       arglag=list(type="integer"))

summary(cb2.rain)

model2 <- glm(Outbreak ~ cb2.rain , family="binomial", Epiweek_avg)

pred.cb2.rain <- crosspred(cb2.rain, model2, from = 0, to = 30, cumul = T)

plot(pred.cb2.rain, "slices", var = 1, col = 3, ci.arg=list(density=15,lwd=2), ci.level=0.80)
plot(pred.cb2.rain, "slices", var = 1, ci = "bars", type="p", col=2, pch=19,
    ci.level=0.80)
plot(pred.cb2.rain, "slices", var = 1, col = 3, cumul = TRUE)

plot(pred.cb2.rain, xlab="Rainfall", zlab="RR", theta=300, phi=30, lphi=30,
    main="3D graph of rainfall effect")

plot(pred.cb2.rain, "contour", xlab="Rainfall", key.title=title("RR"),
    plot.title=title("Contour plot",xlab="Rainfall",ylab="Lag")) 
```



```{r, warnings = FALSE}
cb3.rain <- crossbasis(Epiweek_avg$Total_Rain, lag=12, 
                      argvar=list(fun="poly", degree = 4),
                      arglag=list(fun="integer")
                      )

model3 <- glm(Outbreak ~  cb3.rain, family="binomial", Epiweek_avg)

pred.cb3.rain <- crosspred(cb3.rain, model3, from = 0, to = 30, cumul = T)

plot(pred.cb3.rain, "slices", var = 1, col = 3, ci.arg=list(density=15,lwd=2), ci.level=0.80)
plot(pred.cb3.rain, "slices", var = 1, ci = "bars", type="p", col=2, pch=19,
    ci.level=0.80)
plot(pred.cb3.rain, "slices", var = 1, col = 3, cumul = TRUE)



plot(pred.cb3.rain, xlab="Rainfall", zlab="RR", theta=300, phi=30, lphi=30,
    main="3D graph of rainfall effect")

plot(pred.cb3.rain, "contour", xlab="Rainfall", key.title=title("RR"),
    plot.title=title("Contour plot",xlab="Rainfall",ylab="Lag"))  
```


### CROSS SEC MODEL
```{r}
Data_spring1 <- filter(Epiweek_avg, epiweek == 13 |
                         epiweek == 14 |
                         epiweek == 15 |
                         epiweek == 16 |
                         epiweek == 17)
Data_spring1$Outbreak <- ceiling(Data_spring1$Outbreak)
Data_spring1$wdir <- ifelse(Data_spring1$WDIR < 0, 360 + Data_spring1$WDIR, Data_spring1$WDIR)
Data_spring1$wdir <- as.factor(ifelse(Data_spring1$wdir >= 215 | Data_spring1$wdir <= 35, 1, 0))

crossmodel <- glm(Outbreak ~ Total_Rain + Salinity + PAR  + WSPD + NVcount + wdir, 
               family = "binomial", data = Data_spring1, na.action = na.fail)
summary(crossmodel)
step(crossmodel)
```

```{r}
#Epiweek_avg_test$wdir <- ifelse(Epiweek_avg_test$WDIR < 0, 360 + Epiweek_avg_test$WDIR, Epiweek_avg_test$WDIR)
library(openair)
windRose(
  Epiweek_avg,
  ws = "WSPD",
  wd = "wdir")
hist(Epiweek_avg$wdir, breaks = 100, main = "Wind direction distribution", xlab = "Degrees from north")

#305 and 125
# Perpendicular to above
#215 and 35
```

#Correlation plots
```{r}
library(psych)
library(corrplot)
data <- Epiweek_avg[, c(5,6,9,10,11,12,17)]
data <- na.omit(data)
names(data) <- c("Case Count", "Rainfall", "Salinity", "Sea Temp.", "PAR", "Wind Speed", "Wind Direction")



corrplot.mixed(cor(data),
               lower = "number", 
               upper = "circle",
               tl.col = "black")
```