---
title: "Regression Models with Cumulative Lag Effect"
author: "Marc Wettengel"
date: "3/7/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(dplyr)
library(MASS)
#write.csv(laggeddata, "laggeddata.csv")
lagdata <- read.csv("laggeddata.csv")
lagdata <- lagdata[c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22)]
lagdata$Outbreak <- as.factor(ceiling(lagdata$Outbreak))
names(lagdata)
#lagdataapr <- subset(lagdata, epiweek == 10 | epiweek == 11 |
#                       epiweek == 12 | epiweek == 13 | epiweek == 14 |
#                       epiweek == 15 | epiweek == 16 | epiweek == 17 |
#                       epiweek == 18)


lagdata$WDIR <- ifelse(lagdata$WDIR < 0, 360 + lagdata$WDIR, lagdata$WDIR)
lagdata$wdir <- as.factor(ifelse(lagdata$WDIR >= 215 | lagdata$WDIR <= 35, 1, 0))
lagdata2 <- lagdata
lagdata2 <- na.omit(lagdata2)

```

### ignore this
```{r}
#library(MuMIn)

lagdata2 <- na.omit(lagdata)
modelts <- glm(Outbreak ~ Total_Rain7 + Salinity + PAR + Temperature6 + WSPD7 + NVcount8, 
               family = "binomial", data = lagdata, na.action = na.fail)

summary(modelts)
#exp(modelts$coefficients)
#step(modelts)
#output <- dredge(modelts)
#output
```


# Longitudinal model with random effect
```{r}
library(lme4)
#lagdata$Total_Rain7 <- ifelse(lagdata$Total_Rain7 == 0.83, 3.12, lagdata$Total_Rain7)
lagdata_test <- lagdata
lagdata_test$yearid <- as.factor(lagdata_test$yearid)

glmm1 <- glmer(Outbreak ~ Total_Rain7 + Salinity + PAR + Temperature + WSPD7 + NVcount8 + wdir +(1|yearid), 
               family = binomial(link = "logit"), 
               data = lagdata)
#glmm2 <- glmer(Outbreak ~ Total_Rain7 + Salinity + PAR8 + Temperature6 + WSPD7 + NVcount8 + (1|yearid), 
#               family = binomial(link = "logit"), 
#               data = lagdata)
library(MASS)
glmm2 <- glmmPQL(Outbreak ~ Total_Rain7 + Salinity + PAR + Temperature + WSPD7 + NVcount8 + wdir, random = ~ 1|yearid, 
               family = binomial(link = "logit"), 
               data = lagdata)
# penalized quasi likelihood
summary(glmm1)
summary(glmm2)
#drop1(glmm1)
#qqnorm(glmm1)
#summary(glmm2)
```
# Longitudinal model with random effect but with lag 15 for rainfall
```{r}
glmm3 <- glmmPQL(Outbreak ~ Total_Rain15 + Salinity + Temperature + WSPD7 + NVcount8 + wdir, random = ~ 1|yearid, 
               family = binomial(link = "logit"), 
               data = lagdata2)
# penalized quasi likelihood

summary(glmm3)
glmm3$coefficients
step(glmm3)
glmm4 <- glmer(Outbreak ~ Total_Rain15 + Salinity + Temperature + WSPD7 + NVcount8 + wdir +(1|yearid), 
               family = binomial(link = "logit"), 
               data = lagdata2)
summary(glmm4)
dredge(glmm4)
```


# Longitudinal model without random effect
```{r}
lagdata$yearid <- as.factor(lagdata$yearid)
dat <- groupedData(Outbreak ~ Total_Rain7 + Salinity + PAR8 + Temperature6 + WSPD7 + NVcount8 | yearid, data = lagdata)

lagdata2 <- na.omit(lagdata)
modelts <- glm(Outbreak ~ Total_Rain7 + Salinity + PAR8 + Temperature6 + WSPD7 + NVcount8, 
               family = "binomial", data = lagdata, na.action = na.fail)

summary(modelts)
exp(modelts$coefficients)
step(modelts)
output <- dredge(modelts)
output
```

### GEE
```{r}
library(gee)

fit1 <- gee(Outbreak ~ Total_Rain15 + Salinity + Temperature + WSPD7 + NVcount8 + wdir, id=yearid, data = lagdata2, family = "binomial", corstr = "exchangeable")
summary(fit1)
```





# Mdoel diagnostics:
# fit fitted curve
# Normal QQ plot





# Explain that not all models will hold. Not sure x% of the time the model will hold and how it impacts the final results. This is why we try multiple models and compare results. 

# Spatial temporal model: future work b/c this includes potential spatial effect
# latent characteristic of year is the random effect term. 
# GEE (generalized estimated equation) model as an alternative. Doesn't assume any distribution. 




################################## CROSS-SECTIONAL MODELS

```{r}
crosssecdata <- Epiweek_avg


crosssecdata$WDIR <- as.factor(ifelse(crosssecdata$wdir >= 215 | crosssecdata$wdir <= 35, 1, 0))
crosssecdata <- na.omit(crosssecdata)

crosssecdata2 <- subset(crosssecdata, epiweek == 13 |
                          epiweek == 14 |
                          epiweek == 15 |
                          epiweek == 16 |
                          epiweek == 17 |
                          epiweek == 18 )

```

```{r}
model1 <- glm(Outbreak ~ Total_Rain + Salinity + PAR + Temperature + WSPD + WDIR + NVcount, 
               family = "binomial", data = crosssecdata2, na.action = na.fail)

summary(model1)
exp(model1$coefficients)

model2 <- glm(Outbreak ~ Total_Rain + Salinity + PAR + WSPD + WDIR + NVcount, 
               family = "binomial", data = crosssecdata2, na.action = na.fail)

summary(model2)
exp(model2$coefficients)

model3 <- glm(Outbreak ~ Total_Rain + Salinity + Temperature + WSPD + WDIR + NVcount, 
               family = "binomial", data = crosssecdata2, na.action = na.fail)

summary(model3)
exp(model3$coefficients)
```




