---
title: "Regression Analysis"
author: "Marc Wettengel"
date: "09/01/2023"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
library(ggplot2)
library(MASS)
library(dplyr)

Monthly_avg <- read.csv("Monthly_avg.csv")
Data_complete <- read.csv("Data_complete.csv")
Epiweek_avg <- read.csv("Epiweek_avg.csv")

Monthly_avg$Month <- as.factor(Monthly_avg$Month)
Data_complete$Month <- as.factor(Data_complete$Month)

Monthly_avg$Year <- as.factor(Monthly_avg$Year)
Data_complete$Year <- as.factor(Data_complete$Year)

Monthly_avg$outbreak <- ceiling(Monthly_avg$outbreak)
```

```{r, include = FALSE}
data_log_spring <- subset(Monthly_avg, Month == 3 | Month == 4)
data_log_march <- subset(Monthly_avg, Month == 3)
data_log_apr <- subset(Monthly_avg, Month == 4)
```

```{r, include = FALSE}
monthly_avg <- Monthly_avg[c(1,2,3,4,5,6,7,8,9,10,11,14)]

jan_month <- subset(monthly_avg, Month == 1)
colnames(jan_month)[-1] <- paste0(colnames(jan_month)[-1], "_jan")

feb_month <- subset(monthly_avg, Month == 2)
colnames(feb_month)[-1] <- paste0(colnames(feb_month)[-1], "_feb")

mar_month <- subset(monthly_avg, Month == 3)
colnames(mar_month)[-1] <- paste0(colnames(mar_month)[-1], "_mar")

apr_month <- subset(monthly_avg, Month == 4)
colnames(apr_month)[-1] <- paste0(colnames(apr_month)[-1], "_apr")

data_logistic <- left_join(jan_month, feb_month)
data_logistic <- left_join(data_logistic, mar_month)
data_logistic <- left_join(data_logistic, apr_month)
 

```

# Logistic

### March & April combined
```{r}
log_model1 <- glm(outbreak ~ Total_Rain + Salinity + Temperature + PAR + WSPD,
                  family = 'quasibinomial',
                  data = data_log_spring)
summary(log_model1)

log_model2 <- glm(outbreak ~ Total_Rain + Salinity + Temperature + PAR + WSPD,
                  family = 'binomial',
                  data = data_log_spring)
summary(log_model2)

step(log_model2)
```

### April with March covariates
```{r}
log_model3 <- glm(outbreak_apr ~ Total_Rain_apr + Salinity_apr + Temperature_apr + 
                    PAR_apr + WSPD_apr + Total_Rain_mar + Salinity_mar + Temperature_mar
                  + PAR_mar + WSPD_mar,
                  family = 'binomial',
                  data = data_logistic)
summary(log_model3)
step(log_model3)
```

### April with November covariates
```{r}
# Place holder
# Need to shift November covariates 1 year later
```

### Based off univariate analysis for March & April
```{r}
poisson_model4 <- glm(NVcount ~ Total_Rain + Temperature + PAR, 
                 family = 'poisson', 
                 data = data_poisson)
step(poisson_model4)
```







# Poisson

```{r, include = FALSE}
data_poisson <- subset(Epiweek_avg, epiweek == 10 | epiweek ==11 | epiweek ==12 |
                         epiweek == 13 | epiweek == 14 | epiweek == 15 | 
                         epiweek == 16  | epiweek == 17)
data_poisson <- na.omit(data_poisson)

data_poisson_comp <- subset(Epiweek_avg, epiweek == 1 | epiweek == 2 | 
                         epiweek == 3 | epiweek == 4 | epiweek == 5 | epiweek == 6 |
                           epiweek == 7 | epiweek == 8 | epiweek == 9 | 
                           epiweek == 10 | epiweek ==11 | epiweek ==12 |
                         epiweek == 13 | epiweek == 14 | epiweek == 15 | 
                         epiweek == 16 | epiweek == 17)
data_poisson_comp <- na.omit(data_poisson_comp)

```

### March & April combined
```{r}
poisson_model0 <- glm(NVcount ~ Total_Rain + Salinity + Temperature + PAR + WSPD, 
                 family = 'quasipoisson', 
                 data = data_poisson)
summary(poisson_model0)

poisson_model1 <- glm(NVcount ~ Total_Rain + Salinity + Temperature + PAR + WSPD, 
                 family = 'poisson', 
                 data = data_poisson)

summary(poisson_model1)
step(poisson_model1)
```

### January through April combined
```{r}
poisson_model2 <- glm(NVcount ~ Total_Rain + Salinity + Temperature + PAR + WSPD, 
                 family = 'quasipoisson', 
                 data = data_poisson_comp)

summary(poisson_model2)

poisson_model3 <- glm(NVcount ~ Total_Rain + Salinity + Temperature + PAR + WSPD, 
                 family = 'poisson', 
                 data = data_poisson_comp)

summary(poisson_model3)
step(poisson_model3)
```

### Based off univariate analysis 
```{r}
poisson_model4 <- glm(NVcount ~ Salinity + Temperature + PAR, 
                 family = 'poisson', 
                 data = data_poisson_comp)

summary(poisson_model4)
step(poisson_model4)
```


### Lag model
```{r, include = FALSE}
data_poisson_lag <- subset(Epiweek_avg, epiweek == 1 | epiweek == 2 | 
                         epiweek == 3 | epiweek == 4 | epiweek == 5 | epiweek == 6 |
                           epiweek == 7 | epiweek == 8 | epiweek == 9 | 
                           epiweek == 10 | epiweek ==11 | epiweek ==12 |
                         epiweek == 13 | epiweek == 14 | epiweek == 15 | 
                         epiweek == 16 | epiweek == 17 | epiweek == 53 |
                           epiweek == 52 |epiweek == 51 | epiweek == 50 | 
                           epiweek == 49 | epiweek == 48 | epiweek == 47 |
                           epiweek == 46 | epiweek == 45)
data_poisson_lag <- na.omit(data_poisson_lag)
```

```{r}
poisson_model5 <- glm(NVcount ~ Total_Rain + Salinity + Temperature + PAR + WSPD, 
                 family = 'poisson', 
                 lag = 3,
                 data = data_poisson_lag)
```

























# Summary

- Difference in environmental conditions between outbreaks and non-outbreaks primarily in March, April and November
- Some monthly variation in which environmental condition showed differences, however fairly consistent
- Used t-tests and wilcox test to see if difference is significant 

- Restricted the logistic regression analysis to March and April
    - March & April combined has Wind as a significant predictor
    - April outbreaks with March covariates did not converge except for wind_apr and rain_apr
- November will be included in future rendition 
- Used univariate analysis to select covariates, after then using stepwise selection, Temperature and PAR were significant predictors of outbreak periods
- *No lag has been included 


- For poisson regression, some over dispersion is present (quasipoisson)
- For model selection used regular poisson (stepwise with AIC)
- Did not use march environmental conditions to predict future outbreaks 
    - Epiweeks change & cannot determine when March covariate predicts April case count 
- Lag model is to be created with time frame spanning from November to April to account for earlier environmental conditions

- Based off scatterplots, created a poisson model (PAR, SST and salinity)
    - All of the environmental conditions were significant after stepwise selection
    - Again, no lag was included in this



